Can the symmetric \textit{Fermi} and eROSITA bubbles be produced by tilted jets: a proof-of-concept study

### 2022-03-14 KY-3.14.tex
update

\begin{document}
\title{The symmetry problem of the \textit{Fermi} and eROSITA bubbles: A proof-of-concept study}

@@

\author[0000-0003-3269-4660]{H.-Y. Karen Yang}
\affiliation{Institute of Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan}
\affiliation{Center for Informatics and Computation in Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan}
\affiliation{Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan}

\author[0000-0002-1249-279X]{Hsi-Yu Schive}
\affiliation{Institute of Astrophysics, National Taiwan University, Taipei 10617, Taiwan}
\affiliation{Department of Physics, National Taiwan University, Taipei 10617, Taiwan}
\affiliation{Center for Theoretical Physics, National Taiwan University, Taipei 10617, Taiwan}
\affiliation{Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan}

@@

\begin{abstract}
The \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ on either side of the Galactic center (GC).\
The recent discovery of giant eROSITA bubbles also shows a symmetrically enhanced X-ray\
emission from the shells around the eROSITA bubbles.\
To this end, previous simulation-based works based on active galactic nucleus (AGN) jets model\
assumed that the jets are vertical to the Galactic disk;\
however, there is a lack of preferred orientation of jets with respect to the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic (SRHD) simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium (ISM) disk is an essential\
component for the symmetric \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton (IC) scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
an oblique AGN jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to\
the rotation axis of the Galatic plane (disc normal hereafter).\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles\
composed of turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
has thus released a caveat:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

@@

 \Cref{fig__gammaRay-map} shows the simulated gamma-ray photon flux with a CRe power-law index 2.4\
 compared with the observed one\
 in the energy bin $76.8-153.6$ GeV.\
 As the eROSITA bubbles,\
 one can see that the symmetric \textit{Fermi} bubbles\
 can also be realized by oblique jets.\

@@

>>>>

\begin{document}
%\title{The symmetry problem of the \textit{Fermi} and eROSITA bubbles: A proof-of-concept study}
\title{Can the symmetric \textit{Fermi} and eROSITA bubbles be produced by tilted jets: a proof-of-concept study}

@@

\author[0000-0003-3269-4660]{H.-Y. Karen Yang}
\affiliation{Institute of Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan}
\affiliation{Department of Physics, National Tsing Hua University, Hsinchu 30013, Taiwan}
\affiliation{Center for Informatics and Computation in Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan}
\affiliation{Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan}

\author[0000-0002-1249-279X]{Hsi-Yu Schive}
\affiliation{Institute of Astrophysics, National Taiwan University, Taipei 10617, Taiwan}
\affiliation{Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan}
\affiliation{Department of Physics, National Taiwan University, Taipei 10617, Taiwan}
\affiliation{Center for Theoretical Physics, National Taiwan University, Taipei 10617, Taiwan}

@@

\begin{abstract}
The \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ above and below the Galactic center (GC).\
The recent discovery of giant eROSITA bubbles also shows such a symmetry about the GC, suggesting that they may originate from a single GC activity.\
Previous simulations of bubble formation that invoke active galactic nucleus (AGN) jets\
have assumed that the jets are vertical to the Galactic plane;\
however, in general there does not need to be a correlation between the jet orientation and the rotational axis of the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic (SRHD) simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium (ISM) disk is an essential\
component for the symmetric \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton (IC) scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
an oblique AGN jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to\
the rotation axis of the Galatic plane (disc normal hereafter).\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles\
composed of turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
has thus released a caveat:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

{\color{red} Cite references for pc-scale jet candidates of Sgr A\*, e.g., \citep{Li, Zhu}.}

@@

 \Cref{fig__gammaRay-map} shows the simulated gamma-ray photon flux with a CRe power-law index 2.4\
 compared with the observed one\
 in the energy bin $76.8-153.6$ GeV.\
 As the eROSITA bubbles,\
 one can see that the symmetric \textit{Fermi} bubbles\
 can also be realized by oblique jets.\ {\color{red} KY: Please add brief discussion about (1) the surface of the bubbles, e.g., the simulated surface is not as smooth as the observed one). The instabilities at the bubble surface may be suppressed by the magnetic draping effect (Lyutikov 2006, Y12) if magnetic fields were included in the simulations. (2) the uniform gamma-ray intensity: The simulated gamma-ray intensity distribution shown in Figure 8 does not appear to be as uniform as originally found in Su et al. (2012). As discussed above, the gamma-ray intensity is slightly higher close to the Galactic plane due to the stronger radiation field at lower latitudes. However, this level of brightness variations is consistent with the later observational data of Ackermann et al. (2014) and Selig et al. (2015).}

@@

{\color{red} KY: Please also add a section (either in the Discussion or Appendix) to discuss about the parameter variations. I think the figures you showed me are very important and informative.}

{\color{red} KY: Please add a Discussion section to discuss the cooling times of CR electrons and the requirement for re-acceleration. Emphasize that our modeled gamma-ray and microwave spectra assumed that the underlying CR spectrum is spatially uniform. Therefore our results imply that, whatever the true re-acceleration mechanism is, the CR spectrum has to be spatially uniform with a spectral index of 2.4 in order to fit the observed spectra. Please also include this in the abstract and the conclusions since this is an important point to make.}
